---
title: "Demoing Metadata Class Accuracy"
author: "Clifford Anderson-Bergman"
date: "7/12/2019"
output: html_document
---
This document demos the prediction of metadata. We will demo this with 
the EuCore network. This is an email network of professors at a university. 
```{r}
# Loading libraries and data
library(latChanNet)
library(RcppParallel)

# Edge list pairs. Our model expects node IDs to start with 1, 
# but in this file they start with 0, so we just add 1
edgeList = read.table("email-Eu-core.txt") + 1
# Filtering down to unique undirected edges
edgeList = unq_edges(edgeList)
# For model of Ball, Karrer and Newman (BKN), 
# edge counts are needed. 
# Because simple undirected graph, edge counts are all 1
bkn_edgeList = cbind(edgeList, 1)

# Department for each node
dpt = read.table("email-Eu-core-department-labels.txt")[,2] + 1
# Code expects metadata as data.frame with factors
meta_data = data.frame(dpt = as.factor(dpt))

# Basic data statistics
cat("Number of nodes =", max(edgeList), 
    "\nNumber of edges =", nrow(edgeList), 
    "\nNumber of departments =", length(unique(dpt)), 
    "\nRange of professors per department =", 
    as.numeric(range(table(dpt))))
```
Note that we have some departments that only have a single professor in this dataset!
Of course, we should expect 0 out-of-sample accuracy for nodes that are the only one
in their category.
We will revisit this later. 

Below we have tools for estimating the in- and out-of-sample accuracy.
```{r}
#' @param edgeList edge list
#' @param meta meta data, as data.frame
#' @param K Number of channels
#' @param model Model to use. Options are "LCN" or "BKN"
#' @param nMask Number of metadata variables to mask. Does not mask edges
est_meta_acc = function(edgeList, meta, K = 10, model = "LCN", nMask = 100){
  n_max = max(edgeList)
  meta_name = colnames(meta)
  if(length(meta_name) > 1){
    stop("est_acc currently expects only one column of metadata")
  }
  # Masking subset of metadata
  mask = sample(1:n_max, nMask, replace = F)
  meta_masked = meta
  meta_masked[mask,] = NA
  # Fitting either Latent Channel Model or BKN model
  mod = makeLatentModel(edgeList, nDims = K, 
                        model = model, 
                        metadata = meta_masked)
  res = mod$fit()
  # Extracting most likely prediction for each node
  preds = character(n_max)
  for(i in 1:n_max){
    preds[i] = predict(mod, i, meta_name, type = "metamax")
  }
  ans = NULL
  isCorrect = preds == meta[,meta_name]
  # "in" is in sample accuracy, "out" is out of sample accuracy
  ans["in"] = mean(isCorrect[-mask])
  ans["out"] = mean(isCorrect[mask])
  return(ans)
}

# testing
est_meta_acc(edgeList, meta_data, K = 10, model = "LCN")
est_meta_acc(bkn_edgeList, meta_data, K = 10, model = "BKN")

# Repeat est_meta_acc for better estimation
# Also simplifies input
repeat_acc = function(edgeList, meta, K = 10, MC = 4, 
                      verbose = T){
  lcn_res = NULL
  bkn_res = NULL
  bkn_edges = cbind(edgeList, 1)
  for(i in 1:MC){
    this_lcn_res = est_meta_acc(edgeList, meta, K = K, model = "LCN")
    lcn_res = rbind(lcn_res, this_lcn_res)
    this_bkn_res = est_meta_acc(bkn_edges, meta, K = K, model = "BKN")
    bkn_res = rbind(bkn_res, this_bkn_res)
    if(verbose){
      cat(i, " ")
    }
  }
  if(verbose){
    cat("\n")
  }
  ans = numeric(0)
  ans["lcn_mean_in"] = mean(lcn_res[,"in"])
  ans["lcn_se_in"] = sd(lcn_res[,"in"])/sqrt(MC)
  ans["lcn_mean_out"] = mean(lcn_res[,"out"])
  ans["lcn_se_out"] = sd(lcn_res[,"out"])/sqrt(MC)
  ans["bkn_mean_in"] = mean(bkn_res[,"in"])
  ans["bkn_se_in"] = sd(bkn_res[,"in"])/sqrt(MC)
  ans["bkn_mean_out"] = mean(bkn_res[,"out"])
  ans["bkn_se_out"] = sd(bkn_res[,"out"])/sqrt(MC)
  return(ans) 
}

# testing
repeat_acc(edgeList, meta_data, K = 10, MC = 4)

# Estimate accuracies across several number of channels
mult_acc = function(edgeList, meta_data, K = c(5, 10), MC = 4, verbose = T){
  ans = NULL
  for(k in K){
    if(verbose){
      cat("Starting K =", k, "\n")
    }
    this_acc = repeat_acc(edgeList, meta_data, K = k, MC = MC, verbose = verbose)
    ans = rbind(ans, this_acc)
  }
  rownames(ans) = paste("K =", K)
  return(ans)
}

# Testing
mult_res = mult_acc(edgeList, meta_data)
mult_res

## Plotting results tools
draw_one_res = function(res, x_loc, col, offset){
  points(x_loc, res[1], col = col)
  points(x_loc + offset, res[3], col = col, pch = 16)
  lower = res[1] - 2 * res[2]
  upper = res[1] + 2 * res[2]
  lines(rep(x_loc, 2), c(lower, upper), col = col)
  lower = res[3] - 2 * res[4]
  upper = res[3] + 2 * res[4]
  lines(rep(x_loc + offset, 2), c(lower, upper), col = col)
}

# Plotting mult_acc results
plot_res = function(mult_acc_res, title = "", pnt_offset = 0.25){
  rn = rownames(mult_acc_res)
  Ks = gsub("K = ", "", rn)
  Ks = as.numeric(Ks)
  
  xLim = range(Ks)
  xLim[1] = xLim[1] - pnt_offset * 1.5
  xLim[2] = xLim[2] + pnt_offset * 1.5
  
  plot(NA, xlim = xLim, ylim = c(0,1), 
       main = title, 
       xlab = "Number of Channels", 
       ylab = "Accuracy")

  for(i in 1:nrow(mult_acc_res)){
    this_row = mult_acc_res[i,]
    this_k = Ks[i]
    draw_one_res(this_row[1:4], x_loc = this_k - pnt_offset, 
                 offset = pnt_offset/2, col = "blue")
    draw_one_res(this_row[5:8], x_loc = this_k + pnt_offset, 
                 offset = pnt_offset/2, col = "red")
  }
  legend("bottom", 
         c("LCN: Out of Sample", 
           "LCN: In Sample", 
           "BKN: Out of Sample", 
           "LCN: In Sample"), 
         pch = c(16, 1, 16, 1), 
         col = c("blue", "blue", "red", "red"))
}

# testing
plot_res(mult_res, title = "Testing Code")
```
## Comparing Accuracy on Original Groups
```{r}
K = c(5, 10, 20, 30, 40, 60, 80)
MC = 10

res = mult_acc(edgeList, meta_data, K = K, MC = MC)
plot_res(res, title = "Accuracy on Original Department")
```
# Merging Small Departments
Many of these departments are really small, some on which only have a single professor. We're going to merge this super small departments together to have a more reasonable prediction challenge. 
```{r}
# Any department smaller than this will be merged into "other"
minGrpSize = 15
dpt_merge = dpt
dpt_cnt = table(dpt)
smallDpts = names(dpt_cnt)[dpt_cnt < minGrpSize]
is_small = dpt %in% smallDpts
dpt_merge[is_small] = "other"

# Counts in new categories
table(dpt_merge)

meta_merged = data.frame(dpt = dpt_merge)
merge_res = mult_acc(edgeList, meta_merged, K = K, MC = MC)
plot_res(merge_res, title = "Accuracy on Original Department")
```