---
title: "Analyzing euEmail with Latent Channel Networks"
author: "Clifford Anderson-Bergman"
date: "5/10/2019"
output: md_document
---


For this example, we will analyze a small [email network](https://snap.stanford.edu/data/email-Eu-core.html). 
Nodes in this network are faculty at a European department 
and edges represent emails sent between them. 

First we will load and examine the data. 
```{r}
# Loading edge list
edgeList = read.table("email-Eu-core.txt")
head(edgeList)
```

Note node labels begin with 0. Our code expects nodes to begin with 1, so we will edit this. 

```{r}
edgeList = edgeList + 1
unq_nodes = unique( c(edgeList[,1], edgeList[,2]) ) 
cat("Number of nodes = ", length(unq_nodes),
    "\nNumber of edges = ", nrow(edgeList))
label_table = read.table("email-Eu-core-department-labels.txt")
# Labels don't have to start with 0, but why not
label_table = label_table + 1
# Second column is actual labels, first is ID
head(label_table)
labels = label_table[,2]
```

Note that departments have be anonymized and simply represented as an integer. 

Now we will build and fit our latent channel model.
```{r}
library(latChanNet)

# Number of latent channels
nChannels = 10

# Building (but not fitting) model
# seed is set because initial probabilities randomly assigned
set.seed(123)
mod = makeLCN(edgeList, nChannels)

# Initial log-likelihood
mod$llk()

# Running EM algorithm
res = emLCN(mod, 10000)

# Checking log-likelihood
mod$llk()
```

Now our model has been fit. But how do we get a feel for the structure of the network? One way to do this is to display a heatmap of the fitted probabilities, sorted by department. 

```{r}

# Many of the departments are *really* small, 
# so we are going to bin all departments with 
# less than 15 faculty to be more visually pleasing
minGrpSize = 15

heatmapLCN(mod, labels, minGrpSize = minGrpSize, 
           xlab = "Department Number", 
           ylab = "Channel", 
           main = "Channel Strengths by Department")

```

